### 2021-10-06 Multilevel Missing Data/Multilevel Missing Data.Rmd

---
documentclass: jss
author:
  - name: Hanne Oberman
    affiliation: Utrecht University
    # use this syntax to add text on several lines
    address: |
      | First line
      | Second line
    email: \email{h.i.oberman@uu.nl}
    url: https://hanneoberman.github.io/
  - name: Gerko Vink
    affiliation: 'Utrecht University'
    # To add another line, use \AND at the end of the previous one as above
  - name: Thomas Debray
    # address: |
    #   | Department of Statistics and Mathematics,
    #   | Faculty of Biosciences,
    #   | Universitat Autònoma de Barcelona
    affiliation: |
      | University Medical Centre Utrecht
    # use a different affiliation in adress field (differently formated here)
    affiliation2: Universitat Autònoma de Barcelona
title:
  formatted: "Imputation of Incomplete Multilevel Data with \\pkg{mice}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Imputation of Incomplete Multilevel Data with mice"
  # For running headers, if needed
  short:     "\\pkg{mice}: Multilevel"
abstract: >
  Tutorial paper on imputing incomplete multilevel data with \pkg{mice}. Including methods for ignorable and non-ignorable missingness. AKA missing on so many levels.
keywords:
  # at least one keyword must be supplied
  formatted: [missing data, multilevel, clustering, "\\proglang{R}"]
  plain:     [missing data, multilevel, clustering, R]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
```

# Outline door Thomas

-	Waarom zijn standaard (ad hoc) methoden niet geschikt?

-	Wat is heterogeniteit? (i.e. variability binnen studies vs. variability tussen studies). Random effects voor intercept term, predictor effects, en/of variance residual error

-	JOMO aanpak (misschien buiten beschouwing laten?)

-	MICE aanpak: one-stage imputation versus two-stage imputation methods

-	Imputatie van studie-level versus patient-level covariates

-	Omgaan met MNAR

# References

- [https://onlinelibrary.wiley.com/doi/abs/10.1002/9781119333784.ch18](Chapter: Dealing with Missing Data in an IPD Meta-analysis)

- [https://www.gerkovink.com/miceVignettes/Multi_level/Multi_level_data.html](Vignette: gerkovink.com/miceVignettes/Multi_level)

# Introduction