### 2021-10-11 Multilevel Missing Data/Multilevel-Missing-Data.tex

\documentclass[
]{jss}

\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
Hanne Oberman\\Utrecht University \And Gerko Vink\\Utrecht
University \And Thomas Debray\\University Medical Centre Utrecht
}
\title{Imputation of Incomplete Multilevel Data with \pkg{mice}}

\Plainauthor{Hanne Oberman, Gerko Vink, Thomas Debray}
\Plaintitle{Imputation of Incomplete Multilevel Data with mice}
\Shorttitle{\pkg{mice}: Multilevel}

\Abstract{
Tutorial paper on imputing incomplete multilevel data with \pkg{mice}.
Including methods for ignorable and non-ignorable missingness. AKA
missing on so many levels.
}

\Keywords{missing data, multilevel, clustering, \proglang{R}}
\Plainkeywords{missing data, multilevel, clustering, R}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    Hanne Oberman\\
    Utrecht University\\
    First line\\
Second line\\
  E-mail: \email{h.i.oberman@uu.nl}\\
  URL: \url{https://hanneoberman.github.io/}\\~\\
      }

% Pandoc syntax highlighting

% Pandoc citation processing

\usepackage{amsmath}

\begin{document}

\hypertarget{outline-door-thomas}{%
\section{Outline door Thomas}\label{outline-door-thomas}}

\begin{itemize}
\item
  Why are standard (ad hoc) methods not well suited?
\item
  What is heterogeniety? (i.e.~variability within studies
  vs.~variability between studies). Random effects for intercept term,
  predictor effects, and/or variance residual error
\item
  Systematic vs sporadic missingness
\item
  MICE approach: one-stage imputation versus two-stage imputation
  methods
\item
  Imputation of study level versus patient-level covariates
\item
  Handling MNAR
\item
  Case study: metamisc::impact()
\end{itemize}

\hypertarget{think-about-adding}{%
\section{Think about adding}\label{think-about-adding}}

\begin{itemize}
\tightlist
\item
  JOMO in mice
\end{itemize}

\hypertarget{references}{%
\section{References}\label{references}}

\begin{itemize}
\item
  \href{Chapter:\%20Dealing\%20with\%20Missing\%20Data\%20in\%20an\%20IPD\%20Meta-analysis}{https://onlinelibrary.wiley.com/doi/abs/10.1002/9781119333784.ch18}
\item
  \href{Vignette:\%20gerkovink.com/miceVignettes/Multi_level}{https://www.gerkovink.com/miceVignettes/Multi\_level/Multi\_level\_data.html}
\end{itemize}

\end{document}