### 2021-11-01 Multilevel Missing Data/Multilevel Missing Data.Rmd
Outline paper incl modeling steps

---
documentclass: jss
author:
  - name: Hanne Oberman
    affiliation: Utrecht University
    address: |
      | Padualaan 14
      | 3584 CH Utrecht
    email: \email{h.i.oberman@uu.nl}
    url: https://hanneoberman.github.io/
  - name: Johanna Munoz Avila
    affiliation: University Medical Center Utrecht \AND
  - name: Valentijn de Jong
    affiliation: University Medical Center Utrecht 
  - name: Gerko Vink
    affiliation: Utrecht University \AND
  - name: Thomas Debray
    affiliation: University Medical Center Utrecht 
      # | Julius Center for Health Sciences and Primary Care, 
      # | University Medical Center Utrecht, Utrecht, The Netherlands
    #affiliation2:
title:
  formatted: "Imputation of Incomplete Multilevel Data with \\pkg{mice}"
  # If you use tex in the formatted title, also supply version without
  plain:     "Imputation of Incomplete Multilevel Data with mice"
  # For running headers, if needed
  short:     "\\pkg{mice}: Multilevel"
abstract: >
  Tutorial paper on imputing incomplete multilevel data with \pkg{mice}. Including methods for ignorable and non-ignorable missingness. 
keywords:
  # at least one keyword must be supplied
  formatted: [missing data, multilevel, clustering, "\\pkg{mice}", "\\proglang{R}"]
  plain:     [missing data, multilevel, clustering, mice, R]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
bibliography: ../References/multilevelmice.bib
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
library(tidyverse)

```

# Introduction

## Multilevel data

- What is clustering/multilevel data? In this paper, we discuss grouped observations, not longitudinal data (within-patient clustering).  -> ADD: timeseries also in Discussion section.

- What do we mean by clustering? In the medical field: Clustering by studies (IPDMA), hospitals in registries, multi-center studies etc. In other fields: e.g. official stats clustering at country-level, or social sciences clustering at school-level (related to the sampling design).

- What is heterogeniety? I.e. variability within studies vs. variability between studies

- What does multilevel data look like? ADD: figure to show difference between patient-level datapoints vs cluster-level datapoints. Maybe also add different data frame formats (or just explain in text that there's long and wide formats).

- What methods are required to analyze multilevel data?	Add references, e.g. @hox17. At least explain difference random effects for intercept term, predictor effects, and/or variance residual error.

## Missing data

- Why/where does missingness occur in multilevel data? I.e., not only patient-level but also cluster-level.

- How can we categorize this? Systematic vs sporadic missingness, see @resc13. ADD: visualization of systematic vs sporadic missingness. Within systematic we have always missing (same value per cluster) and non-measured variables (may differ per patient). TODO: adjust md pattern to match text. -> syst may vary or same for all patients (observations/participants).

```{r patterns, echo=FALSE}
# monotone vs non-monotone
patterns <- rbind(
  tibble(pat = rep("Systematic", 5), rownr = 5:1, X = rep(1, 5), Y = c(1,1,1,0, 0), Z = c(1, rep(0,4))) %>% pivot_longer(cols = c(X, Y, Z)),
  tibble(pat = rep("Sporadic", 5), rownr = 5:1, X = c(0, rep(1,4)), Y = c(1,1,1,0, 0), Z = c(1, rep(0,3), 1)) %>% pivot_longer(cols = c(X, Y, Z))) %>%
  mutate(across(everything(), as.factor))
ggplot(patterns, aes(x = name, y = rownr, fill = value, type = pat)) +
  geom_tile(color = "black") +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(~pat) +
  theme_void() +
  theme(legend.position = "none") 
```

-	ADD: missingness mech here

- Why are standard (ad hoc) missing data methods not well suited?

- What types of multilevel methods are available? General overview of approaches, see @audi18 @grun18. E.g., imputation of study level versus patient-level covariates, and one-stage imputation versus two-stage imputation methods.

## Aim of this paper

- Provide practical guidelines with code snippets for imputation of incomplete multilevel data.

- We focus on the workflow for conditional modeling (not JOMO) in `mice`. Refer to other packages: `mitml`, `miceadds`.

- Case study options: `metamisc::impact` (real data on traumatic brain injuries, IPD), `mice::popularity` (simulated data with MNAR/MAR mixture, schools). -> Check Gelman's data/NSRI data.

# Workflows

## Case study

- Introduce case study and set scope of this tutorial: We're providing an overview of implementations. It's up-to the reader to decide which strategy suits their data. So we won't go into detail for the different methods (and equations). This paper is just a software tutorial. We'll keep it practical. -> ADD: some kind of help function that suggests a suitable pred matrix to the user, given a certain analysis model.

## Modeling choices

- Which models will we discuss? We'll build the model to grow in complexity. The final model is the most complex but also the most versatile.

- Note on model complexity: Typically, we should at least use random intercepts, but often random slopes as well. Ideally we impute with random everything and heteroscedastic errors: most generic method (no worry about congeniality, but don't mention the term) ->  Refer to other papers for background, we'll focus just on the software implementation of the situations mentioned there. Sometimes there's little reason to assume some variable is affected by heterogeneity. -> Refer to Meng, Vincent, and a paper by Grund on congeniality and random slopes.

- Step 0: study as a predictor, AKA multilevel imputation for dummies. Doesn't work for syst missing.

## Conditional models

- How to define the imputation model(s) in `mice`?

- What do the different implementations look like?

- Step 1: Intercept

- Step 2: Slope

- Step 3: Residuals

- Heckman model for MNAR

## Pooling

- Analysis of scientific interest.

- Pooling using `mitml`.

- Pooling 'regular' parameters vs more 'exotic' parameters (SE of residual errors, or autocorreletion)

- ADD: export mids objects to other packages like `lme4` or `coxme`(?)

# Discussion

- JOMO in `mice` --> on the side for now

- Additional levels of clustering

- Timeseries: and polynomial relationship in the clustering.

# References